A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2018-06-07 paper/paper.tex
save nodes over 100
Starting a bit on the experiments.

% ============================================================
\begin{document}
\maketitle
\linenumbers
% ==================================================================

@@

\section{Introduction}
\label{sec:introduction}

@@

\section{Related Work}
\label{sec:relatedwork}

@@

\section{Description of our Heuristic Approach}
\label{sec:algorithm}

@@

\myparagraph{Avoiding local maxima}
%
To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$. The first mimics the human behaviour. What would~one do to escape from a locally optimal solution? She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that 
%by changing the embedding of the graph or its crossing structure, 
by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows: 
%
\begin{inparaenum}[(i)]
%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices. 
\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} containing all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
\item afterwards, the vertex-pool switches back to the critical vertices.
\end{inparaenum}
%
%From our extensive practical analysis, we quickly realized that, while 
While this approach turned out to be quite effective for medium-size graphs, for larger graphs, unfortunately, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.

@@

\myparagraph{Complexity related issues}
%
A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing). Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.

@@

\subsection{Some interesting variants}

@@

\myparagraph{Aspect ratio}
%
Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings. It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value. What we simply had to do was to reject candidate positions, which violate this precondition.

\myparagraph{Total resolution} 
%
The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}. Formally, it is defined as the minimum of these two measures. It was not difficult to adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing. In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution. Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease. We do so by simply rejecting candidate positions, which yield a reduced total resolution.

\myparagraph{Angular resolution} 
%  
As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution). We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).

\myparagraph{Grid drawings}
%
Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates. However, it can be easily adjusted to produce such drawings. More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid). One can even bound the size of the grid, which simply implies that candidate grid positions outside the bounds must be rejected.

@@

\section{Experimental Evaluation}
\label{sec:experiments}

\begin{itemize}
\item Experimental setup.
\item Benchmarks + degree-3.
\item Aesthetics:

\begin{itemize}
\item Number of crossings (\# of crossings).
\item Average size of crossing angles (angle size).
\item Average of edge lengths (edge length).
\item Angular resolution (angular res.).
\end{itemize}

\item Results.
\item Discussion.
\end{itemize}

\section{Conclusions}
\label{sec:conclusions}

>>>>

% ==================================================================
\begin{document}
\maketitle
\linenumbers
% ==================================================================

@@

% ==================================================================
\section{Introduction}
\label{sec:introduction}
% ==================================================================

@@

% ==================================================================
\section{Related Work}
\label{sec:relatedwork}
% ==================================================================

@@

% ==================================================================
\section{Description of our Heuristic Approach}
\label{sec:algorithm}
% ==================================================================

@@

% ==================================================================
\myparagraph{Avoiding local maxima}
% ==================================================================
%
To avoid getting trapped to locally optimal solutions, we mainly investigated two approaches, which are both  parametrizable by two input parameters $\zeta$ and $\zeta'$. The first mimics the human behaviour. What would~one do to escape from a locally optimal solution? She would stop trying to move the endvertices of the edges defining the crossing resolution; she would rather start moving ``irrelevant'' vertices hoping that 
%by changing the embedding of the graph or its crossing structure, 
by doing so a better solution will be easier to be computed afterwards. Our algorithm is mimicking this idea as follows: 
%
\begin{inparaenum}[(i)]
%\item As long as the algorithm detects that within the last $\zeta$ iterations the crossing resolution of the graph has being improved, the vertex-pool contains only critical vertices. 
\item if during the last $\zeta$ iterations the crossing resolution has not been~improved, then the vertex-pool becomes \emph{wider} containing all the vertices, and the algorithm is executed with this vertex-pool for $\zeta'$ iterations;
\item afterwards, the vertex-pool switches back to the critical vertices.
\end{inparaenum}
%
%From our extensive practical analysis, we quickly realized that, while 
While this approach turned out to be quite effective for medium-size graphs, for larger graphs, unfortunately, it was not so efficient; in most iterations with the wider vertex-pool, the embedding could not change in a beneficial way for the algorithm~to~proceed.

@@

% ==================================================================
\myparagraph{Complexity related issues}
% ==================================================================
%
A critical factor that highly affects the efficiency of our algorithm is the computation of the crossing points of the edges (and the corresponding angles at these points, which determine the crossing resolution of the drawing). Given  a drawing, a naive approach to compute its crossings requires $O(m^2)$ time, which can be slightly improved using a standard plane-sweep technique to $O(m \log m + c)$ time, where $m$ and $c$ are the number of edges and crossings of the drawing, respectively; see, e.g.,~\cite{DBLP:books/lib/BergCKO08}.

@@

% ==================================================================
\subsection{Some interesting variants}
\label{ssec:variants}
% ==================================================================

@@

% ==================================================================
\myparagraph{Aspect ratio}
% ==================================================================
%
Formally, the aspect ratio of a drawing is the ratio of the length of its longest edge to the length of its shortest edge. Sometimes it is also used as a measure of the area of non-grid drawings. It was easy to instruct our algorithm to prevent producing drawings with aspect ratio either higher than the one of the starting layout or higher than a given input value. What we simply had to do was to reject candidate positions, which violate this precondition.

% ==================================================================
\myparagraph{Total resolution} 
% ==================================================================
%
The notion of the total resolution of a drawing was introduced relatively recently with aim of ``balancing'' the measures of the crossing and of the angular resolution of a drawing~\cite{DBLP:journals/cj/ArgyriouBS13}. Formally, it is defined as the minimum of these two measures. It was not difficult to adjust our algorithm to yield drawings with high total resolution by simply taking into account also the angular resolution of the drawing. In particular, if the total resolution of the drawing is defined by its angular resolution, then the way we compute the critical vertices of this drawing has to change; the critical vertices must be the endvertices of the pairs of edges that define the angular resolution. Also, at each iteration of our algorithm we have to ensure that the total resolution does not decrease. We do so by simply rejecting candidate positions, which yield a reduced total resolution.

% ==================================================================
\myparagraph{Angular resolution} 
% ==================================================================
%  
As it is the case with the force-directed algorithms of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} and Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, our algorithm can be also restricted to maximize only the angular resolution (by neglecting its crossing resolution). We already described in the previous paragraph the necessary changes in the definition of the critical vertices and the rule according to which a candidate position is rejected (i.e., when it yields a drawing with a reduced angular resolution).

% ==================================================================
\myparagraph{Grid drawings}
% ==================================================================
%
Our algorithm, as it has been described so far, does not necessarily produce grid drawings, i.e., drawings in which the vertices are at integer coordinates. However, it can be easily adjusted to produce such drawings. More precisely, if we round the candidate positions computed at each iteration of our algorithm to their closest grid points and use these grid points as candidates for the next position of the vertex to be moved, then the obtained drawing will be grid (assuming, of course, that the starting drawing is grid). One can even bound the size of the grid, which simply implies that candidate grid positions outside the bounds must be rejected.

@@

% ==================================================================
\section{Experimental Evaluation}
\label{sec:experiments}
% ==================================================================

In this section, we present the results of our experimental evaluation. For comparison purposes, apart from our algorithm, we have also implemented the force-directed algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}. The implementations were in Java using the yFiles library~\cite{DBLP:books/sp/04/WieseE004}. The experiment was performed on a Linux laptop with four cores at 2.5 GHz and 3 GB RAM.
%
As a test set for our experiment, we used the non-planar Rome graphs~\cite{DBLP:reference/crc/BattistaD13}, which form a collection of around 8.100 benchmark graphs (commonly used for testing the efficiency of algorithms for drawing graphs).

The experiment was performed as follows. Initially, each Rome graph was laid out using the SmartOrganic layouter of yFiles~\cite{DBLP:books/sp/04/WieseE004}. Starting from this layout, every graph was drawn with 
\begin{inparaenum}[(i)]
\item our algorithm, 
\item our algorithm adjusted not to violate the aspect ratio of the initial layout, 
\item the force-directed algorithm of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13}, and
\item the force-directed algorithm of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13}.
\end{inparaenum}
%
We compared the quality of the produced drawings based on the following aesthetic properties:
%
\begin{enumerate}[P.1.]
\item \label{p:cr-res} crossing resolution, 
\item \label{p:to-res} total resolution, 
\item \label{p:an-res} angular resolution, 
\item \label{p:as-rat} aspect ratio, and
\item \label{p:no-xing} number of crossings.
\end{enumerate}

The comparisons were on average across dirrerent drawings with same number of vertices. For Properties~P.\ref*{p:cr-res}, P.\ref*{p:to-res} and P.\ref*{p:an-res}, we further adjusted each of the algorithms of the experiment to optimize the corresponding measures. Then, for each of them we reported the average of the corresponding aspect ratios and number of crossings of the produced layouts. Note that all algorithms of the experiment can easily be adjusted to maximize only the crossing resolution, or only the angular resolution or both (by maximizing the total resolution). In our algorithm, this can be achieved by adjusting appropriately the content of the vertex-pool (as we saw in Section~\ref{ssec:variants}), while in the algorithms of Argyriou et al.~\cite{DBLP:journals/cj/ArgyriouBS13} and of Huang et al.~\cite{DBLP:journals/vlc/HuangEHL13} by appropriately selecting the set of forces to use (as each of them, has a different set of forces to maximize the crossing and the angular resolution, such that together they maximize the total resolution)\todo{up to here}.

. If both the angular and the
crossing resolution between two consecutive iterations of each
algorithm were not improved more than 0.001 degrees, we assumed that
the algorithm has converged and we did not proceed any more. We note
that the termination condition is quite strict and demands a large
number of iterations. The maximum number of iterations that an
algorithm could perform in order to converge was set to 100.000.
However, very few graphs required the maximum number of iterations
in order to converge to a layout. Once an algorithm converged, we
measured its angular, crossing and total resolution, its average
edge length, the number of crossings, the running time, and the
required number of iterations to converge.

\begin{itemize}
\item Results.
\item Discussion.
\end{itemize}

% ==================================================================
\section{Conclusions}
\label{sec:conclusions}
% ==================================================================
